Distance Sampling in \proglang{R}

### 2017-11-05 paper.Rnw
EAR comments

The detection function models the probability $\Prob(\text{object detected } \vert \text{ object at distance } y)$ and is usually denoted $g(y; \boldsymbol{\theta})$ where $y$ is distance (from a line or point) and $\boldsymbol{\theta}$ is a vector of parameters to be estimated. The parameter(s) $\boldsymbol{\theta}$ are used in a detection function (see Section 3.1), serving as the basis to produce an estimate of the probability that an object within the truncation distance of the point or line transect is detected during the survey. Our goal is to estimate an \textit{average probability of detection} ($p$, average in the sense of an average over distance from $0$ to truncation distance $w$), so we must integrate out distance from the detection function. Letting $x$ denote a perpendicular distance from a line and $r$ denote radial distance from a point:
$$
p = \left\{
\begin{array}{l l}
\int_0^w \frac{1}{w}    g(x; \boldsymbol{\theta}) dx & \quad \text{for lines,} \\
\int_0^w \frac{2r}{w^2} g(r; \boldsymbol{\theta}) dr & \quad \text{for points,}
\end{array} \right.
$$
where the fractions pre-multiplying the detection function describes the distribution of objects with respect to the sampler, taking into account the geometry of the sampler \citep[usually referred to as the \textit{probability density function of (object) distances} and denoted $\pi(y)$;][Chapter 3]{Buckland:2001vm}. Figure \ref{fig:detectdists} shows the relationship between the detection function, the probability density function of object distances and the probability density function of observed distances.

@@

It is not always necessary to include adjustments (except in the case of the uniform key) and in such cases we refer to these as ``key only'' models (see the next section and ``Model checking and model selection''). Adjustment terms increase the flexibility of the detection functioni but this added flexibility comes at the expense of additional parameters to be estimated. The analyst must consider whether the process giving rise to the distribution of detection distances should be modelled with shapes depicted in Figure \ref{fig:keyadj}.

@@

1. \code{Summary statistics}: giving the areas, covered areas, effort, number of observations, number of transects, encounter rate, its standard error and coefficient of variation for each stratum. 
2. \code{Abundance}: giving estimates, standard errors, coefficients of variation, lower and upper confidence intervals and finally the degrees of freedom for each stratum's abundance estimate.
3. \code{Density}: lists the same statistics as \code{Abundance} but for a density estimate.

>>>>

The detection function models the probability $\Prob(\text{object detected } \vert \text{ object at distance } y)$ and is usually denoted $g(y; \boldsymbol{\theta})$ where $y$ is distance (from a line or point) and $\boldsymbol{\theta}$ is a vector of parameters to be estimated. The parameter(s) $\boldsymbol{\theta}$ are used in a detection function (see Section 3.1), serving as the basis to produce an estimate of the probability that an object within the truncation distance of the point or line transect is detected during the survey. Our goal is to estimate an \textit{average probability of detection} ($p$, average in the sense of an average over distance from $0$ to truncation distance $w$), so we must integrate out distance from the detection function. Letting $x$ denote a perpendicular distance from a line and $r$ denote radial distance from a point:
$$
p = \left\{
\begin{array}{l l}
\int_0^w \frac{1}{w}    g(x; \boldsymbol{\theta}) dx & \quad \text{for lines,} \\
\int_0^w \frac{2r}{w^2} g(r; \boldsymbol{\theta}) dr & \quad \text{for points,}
\end{array} \right.
$$
where the fractions pre-multiplying the detection function describes the distribution of objects with respect to the sampler, taking into account the geometry of the sampler \citep[usually referred to as the \textit{probability density function of (object) distances} and denoted $\pi(y)$;][Chapter 3]{Buckland:2001vm}. Figure \ref{fig:detectiondist} shows the relationship between the detection function, the probability density function of object distances and the probability density function of observed distances.

@@

It is not always necessary to include adjustments (except in the case of the uniform key) and in such cases we refer to these as ``key only'' models (see the next section and ``Model checking and model selection''). Adjustment terms increase the flexibility of the detection function but this added flexibility comes at the expense of additional parameters to be estimated. The analyst must consider whether the process giving rise to the distribution of detection distances should be modelled with shapes depicted in Figure \ref{fig:keyadj}.

@@

\begin{enumerate}
\item \code{Summary statistics}: giving the areas, covered areas, effort, number of observations, number of transects, encounter rate, its standard error and coefficient of variation for each stratum. 
\item \code{Abundance}: giving estimates, standard errors, coefficients of variation, lower and upper confidence intervals and finally the degrees of freedom for each stratum's abundance estimate.
\item \code{Density}: lists the same statistics as \code{Abundance} but for a density estimate.
\end{enumerate}
